### 2016-06-15 Paper_v4.tex
Many updates

Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}C).
A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lane 2).
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
even 10\% of the nominal C3 was able to generated enough C5 convertase to maintain C5a formation.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).

@@

In this study, we developed an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
The only intervention that consistently reduced C5a formation for all cases was a knockdown of both C3 and C5.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.

@@

Global analysis of the complement model suggested potentially important therapeutic targets.
Complement malfunctions are implicated in a spectrum of diseases, however the
development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
In this study, we analyzed a validated ensemble of reduced order complement models to better understand the strengths and weaknesses of the cascade.
In the presence of an initiator, C3a and C5a formation was sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
For example, Omeros, a protease inhibitor that targets the MASP-2 complex, has been shown to inhibit the formation of downstream convertases \cite{schwaeble2011methods}.
Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
Removing supporting molecules could also destabilize the convertases.
For example, Novelmed Therapeutics developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
The most commonly used anti-complement drug eculizumab, targets the C5 protein \cite{morgan2015complement}.
A number of other drugs targeting C5 are also being developed. For example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
while Mubodina, an antibody that targets C5, is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
Other agents such as the small molecule Coversin \cite{weston2014clinical}, the aptamer Zimura \cite{epstein2007complement},
as well as small peptides and RNAi could also be used to inhibit C5 \cite{borodovsky2014aln}.
The cyclic peptide inhibitor Compstatin and its derivatives, which target C3, are promising approaches for the treatment of X, Y and Z [FINISH] \cite{Mastellos:2015aa}.
However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
For example, the eculizumab dosage must be significantly adjusted during the course of treatment for aHUS \cite{noris2014dynamics}.
A validated complement model, in combination with personalized pharmacokinetic models of immune system function, could be an important development for the field.

@@

The performance of the complement model was impressive given its limited size.
However, there are several questions that should be explored further.
A logical progression for this work would be to expand the network to include the classical pathway and the formation of the membrane attack complex (MAC).
However, time course measurements of MAC abundance (and MAC formation dynamics) are scarce, making the inclusion of MAC challenging.
On the other hand, inclusion of classical pathway activation is straightforward.
Liu et al., have shown cross-talk between the activation of the classical and lectin pathways through C reactive proteins (CRP) and L-ficolin (LF) under inflammation conditions \cite{liu2011computational}. Thus, inclusion of these species, in addition to a lumped activation term for the classical pathway should allow us to capture classical activation.
Next, we should address the C3a time scale issue.
We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
Tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
Thus, we skipped an important step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
The model should be expanded to include the C3wBb intermediate, where C3wBb catalyzes C3 cleavage at a slow rate compared to normal AP or CP C3 convertases.
We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
This shortcoming could be addressed by including balances around factor B and D, and including these species in the appropriate kinetic rates.
The C5a predictions also had an accelerated time scale.
However, because the C5a time scale depended strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
Lastly, we should also consider including the C2-by pass pathway, which was not included in the model.
The C2-bypass mediates lectin pathway activation, without the involvement of MASP-1/2.
Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement model performance.

@@

\subsection*{Estimating an ensemble of complement model parameters.}
We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique [REF].
DOPS is a novel hybrid meta-heuristic which combines a multi-swarm particle swarm method with the dynamically dimensioned search approach of Shoemaker and colleagues \cite{Tolson2007}.
DOPS minimized the squared residual between simulated and C3a and C5a measurements with and without zymosan as a single objective.
The best fit set estimated by DOPS served as the starting point for multiobjective ensemble generation using Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs) \cite{JuPOETs-BioArXiv}.
JuPOETs is a multiobjective approach which integrates simulated annealing with Pareto optimality to estimate model ensembles on or near the optimal tradeoff surface between competing training objectives. JuPOETs minimized training objectives of the form:
\begin{equation}\label{eqn_cost2}
	O_{j}(\mathbf{k}) = \sum_{i=1}^{\mathcal{T}_{j}}\biggl(\hat{\mathcal{M}}_{ij}-\hat{y}_{ij}(\mathbf{k})\biggr)^2 + \left(\frac{{\mathcal{M}^{\prime}_{ij}}-\max{y_{ij}}}{{\mathcal{M}^{\prime}_{ij}}}\right)^{2}
\end{equation}subject to the model equations, initial conditions and parameter bounds $\mathcal{L}\leq\mathbf{k}\leq\mathcal{U}$.
The first term in the objective function measured the shape difference between the simulations and measurements.
The symbol $\hat{\mathcal{M}}_{ij}$ denotes a scaled experimental observation (from training set $j$) while the symbol $\hat{y}_{ij}$ denotes the scaled simulation output (from training set $j$). The quantity $i$ denotes the sampled time-index and $\mathcal{T}_{j}$ denotes the number of time points for experiment $j$.
The scaled measurement is given by:
\begin{equation}\label{norm_exp_data}
\hat{\mathcal{M}}_{ij} = \frac{\mathcal{M}_{ij} - \min_{i}\mathcal{M}_{ij}}{\max_{i}{\mathcal{M}_{ij}}-\min_{i}{\mathcal{M}_{ij}}}
\end{equation}
Under this scaling, the lowest measured concentration become zero while the highest equaled one, where a similar scaling was defined for the simulation output.
The second-term in the objective function quantified the absolute error in the estimated concentration scale, where the absolute measured
concentration (denoted by $\mathcal{M}^{\prime}_{ij}$) was compared with the largest simulated value.
In this study, we minimized two training objectives, the total C3a and C5a residual w/o zymosan (O$_{1}$) and the total C3a and C5a residual for 1 mg/ml zymosan (O$_{2}$).
JuPOETs identified an ensemble of N $\simeq$ 2100 parameter sets which were used for model simulations and uncertainty quantification subsequently.
JuPOETs is open source, available under an MIT software license.
The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
The objective functions used in this study are available in the GitHub model repository available from http://varnerlab.org.

>>>>

Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lane 2).
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
even 10\% of the nominal C3 was able to generated enough C5 convertase to maintain C5a formation.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).

@@

In this study, we developed an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
The only intervention that consistently reduced C3a and C5a formation for all cases was a knockdown of both C3 and C5.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.

@@

Global analysis of the complement model suggested potentially important therapeutic targets.
Complement malfunctions are implicated in a spectrum of diseases, however the
development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
In this study, we analyzed a validated ensemble of reduced order complement models to better understand the strengths and weaknesses of the cascade.
In the presence of an initiator, C3a and C5a formation was sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
For example, Omeros, a protease inhibitor that targets the MASP-2 complex, has been shown to inhibit the formation of downstream convertases \cite{schwaeble2011methods}.
Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
Removing supporting molecules could also destabilize the convertases.
For example, Novelmed Therapeutics developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
The most commonly used anti-complement drug Eculizumab, targets the C5 protein \cite{morgan2015complement}.
Several other antibodies targeting C5 are also being developed; for example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
while Mubodina is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
Other agents such as Coversin \cite{weston2014clinical} or the aptamer Zimura \cite{epstein2007complement} could also be used to knockdown C5.
The cyclic peptide inhibitor Compstatin and its derivatives, which target C3, are promising approaches for the treatment of X, Y and Z [FINISH] \cite{Mastellos:2015aa}.
However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
For example, the eculizumab dosage must be significantly adjusted during the course of treatment for aHUS \cite{noris2014dynamics}.
A validated complement model, in combination with personalized pharmacokinetic models of immune system function, could be an important development for the field.

@@

The performance of the complement model was impressive given its limited size.
However, there are several questions that should be explored further.
A logical progression for this work would be to expand the network to include the classical pathway and the formation of the membrane attack complex (MAC).
However, time course measurements of MAC abundance (and MAC formation dynamics) are scarce, making the inclusion of MAC challenging.
On the other hand, inclusion of classical pathway activation is straightforward.
Liu et al., have shown cross-talk between the activation of the classical and lectin pathways through C reactive proteins (CRP) and L-ficolin (LF) under inflammation conditions \cite{liu2011computational}. Thus, inclusion of these species, in addition to a lumped activation term for the classical pathway should allow us to capture classical activation.
Next, we should address the C3a time scale issue.
We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
Tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
Thus, we skipped an important step which could strongly influence AP C3 convertase formation by slowing down the rate C3 cleavage into C3a and C3b.
The model should be expanded to include the C3wBb intermediate, where C3wBb catalyzes C3 cleavage at a slow rate compared to normal AP or CP C3 convertases.
We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
This shortcoming could be addressed by including balances around factor B and D, and including these species in the appropriate kinetic rates.
The C5a predictions also had an accelerated time scale.
However, because the C5a time scale depended strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
Lastly, we should also consider including the C2-by pass pathway, which was not included in the model.
The C2-bypass mediates lectin pathway activation, without the involvement of MASP-1/2.
Thus, inclusion of this pathway could be important for understanding the role of MASP-1/2 inhibitors on complement activation.

@@

\subsection*{Estimating an ensemble of complement model parameters.}
We estimated a single initial parameter set using the Dynamic Optimization with Particle Swarms (DOPS) technique \cite{DOPS2016}.
DOPS is a novel hybrid meta-heuristic which combines a multi-swarm particle swarm method with the dynamically dimensioned search approach of Shoemaker and colleagues \cite{Tolson2007}.
DOPS minimized the squared residual between simulated and C3a and C5a measurements with and without zymosan as a single objective.
The best fit set estimated by DOPS served as the starting point for multiobjective ensemble generation using Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs) \cite{JuPOETs-BioArXiv}.
JuPOETs is a multiobjective approach which integrates simulated annealing with Pareto optimality to estimate model ensembles on or near the optimal tradeoff surface between competing training objectives. JuPOETs minimized training objectives of the form:
\begin{equation}\label{eqn_cost2}
	O_{j}(\mathbf{k}) = \sum_{i=1}^{\mathcal{T}_{j}}\biggl(\hat{\mathcal{M}}_{ij}-\hat{y}_{ij}(\mathbf{k})\biggr)^2 + \left(\frac{{\mathcal{M}^{\prime}_{ij}}-\max{y_{ij}}}{{\mathcal{M}^{\prime}_{ij}}}\right)^{2}
\end{equation}subject to the model equations, initial conditions and parameter bounds $\mathcal{L}\leq\mathbf{k}\leq\mathcal{U}$.
The first term in the objective function measured the shape difference between the simulations and measurements.
The symbol $\hat{\mathcal{M}}_{ij}$ denotes a scaled experimental observation (from training set $j$) while the symbol $\hat{y}_{ij}$ denotes the scaled simulation output (from training set $j$). The quantity $i$ denotes the sampled time-index and $\mathcal{T}_{j}$ denotes the number of time points for experiment $j$.
The scaled measurement is given by:
\begin{equation}\label{norm_exp_data}
\hat{\mathcal{M}}_{ij} = \frac{\mathcal{M}_{ij} - \min_{i}\mathcal{M}_{ij}}{\max_{i}{\mathcal{M}_{ij}}-\min_{i}{\mathcal{M}_{ij}}}
\end{equation}
Under this scaling, the lowest measured concentration become zero while the highest equaled one, where a similar scaling was defined for the simulation output.
The second-term in the objective function quantified the absolute error in the estimated concentration scale, where the absolute measured
concentration (denoted by $\mathcal{M}^{\prime}_{ij}$) was compared with the largest simulated value.
In this study, we minimized two training objectives, the total C3a and C5a residual w/o zymosan (O$_{1}$) and the total C3a and C5a residual for 1 mg/ml zymosan (O$_{2}$).
JuPOETs identified an ensemble of N $\simeq$ 2100 parameter sets which were used for model simulations and uncertainty quantification subsequently.
JuPOETs is open source, available under an MIT software license.
The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
The objective functions used in this study are available in the GitHub model repository available from http://varnerlab.org.
